---
title: "Foo"
author: "Bar"
date: "10/07/2014"
output: html_document
---


- Union: 2.2 million (about 2 million whites, 200,000 blacks) (Vinovskis (1989), p. 40).
- Confederate: 0.9 million  (Vinovskis (1989), p. 40)

```{r}
library("dplyr")
library("lubridate")
library("rstan")
library("ggplot2")
library("reshape2")

datapath <- "~/Documents/research/acward/data/greer2005/"
datafile <- "greer2005_weekly_casualties.csv"

fill_na <- function(x, fill = 0) {
  x[is.na(x)] <- fill
  x
}

## Beta distribution method of moments
beta_mm <- function(xbar, vbar) {
  xx <- xbar * (1 - xbar)
  a <- xbar * (xx / vbar - 1)
  b <- (1 - xbar) * (xx / vbar - 1)
  list(a = a, b = b)
}

casualties <- 
  (read.csv(file.path(datapath, datafile),
                      stringsAsFactors = FALSE)
   %>% mutate(date = as.Date(date))
   )

casualties_week <- 
  (group_by(casualties, date)
   %>% mutate(confederate = fill_na(confederate),
              union = fill_na(union))
   %>% filter(battle != "attrition")
   %>% summarise(confederate = sum(confederate),
                 union = sum(union))
   %>% merge(data.frame(date = seq(as.Date("1861-07-07"), as.Date("1865-06-18"), by = "week")),
             all.x = TRUE)
   %>% mutate(confederate = fill_na(confederate),
              union = fill_na(union))
   )

```

# Mexican American War Data

```{r}
mex_war <-
  (read.csv("data/Mexican-American-War-Battles.csv", stringsAsFactors = FALSE)
   %>% mutate(start_date = as.Date(start_date, "%m/%d/%Y"),
              end_date = as.Date(end_date, "%m/%d/%Y"))
  )
mex_war_weeks <-
  data.frame(date = seq(as.Date("1846-4-25"), as.Date("1848-2-2"), by = "weeks"))
mex_war_weeks_days <-
  (mex_war_weeks
   %>% rowwise() %>% do(data.frame(week_date = .$date, date = .$date + 0:6))                                    
   )
mex_war_weeks <-
  (merge(mex_war, mex_war_weeks_days, 
         by.x = "end_date", by.y = "date")
   %>% group_by(week_date)
   %>% summarise(MEX = sum(MEX), USA = sum(USA))
   %>% merge(mex_war_weeks, all.y = TRUE,
             by.x = "week_date", by.y = "date")
   %>% mutate(MEX = fill_na(MEX), USA = fill_na(USA))
   %>% plyr::rename(c(week_date = "date"))
   )
```


# Casualty distribution

```{r}
(ggplot(melt(casualties_week, id.vars = "date"), aes(x = date, y = value, color = variable))
 + geom_line())
```

Plot of the distributions of weekly casualties:

```{r}
(ggplot(melt(casualties_week, id.vars = "date"), aes(x = value, color = variable))
 + geom_density() + geom_rug())

```

What are the parameters of these distributions (assuming log-normal distribution)

```{r}
(melt(casualties_week, id.vars = "date")
 %>% group_by(variable)
 %>% summarise(casualties_mean = mean(log(value + 1)),
               casualties_sd = sd(log(value + 1))
               )
 )
```

Constant model
```{r}
m1 <- stan_model("stan/model1.stan")
```

```{r}
mex_war_data <- within(list(), {
  y <- log(mex_war_weeks[ , c("USA", "MEX")] + 1)
  n <- nrow(y)
  mu_mean <- apply(y, 2, mean)
  mu_cov <- diag(rep(1e6, 2))
  sigma_mean <- apply(y, 2, sd)
  sigma_var <- rep(1e6, 2)
  rho_a <- 1
  rho_b <- 1
})            
```

Mexican war prior
```{r}
res_mex_war <- sampling(m1, data = mex_war_data)
res_mex_war_summary <- summary(res_mex_war)[[1]]
res_mex_war_summary_rho <- beta_mm(res_mex_war_summary["rho", "mean"],
                                    res_mex_war_summary["rho", "sd"] ^ 2)

mex_war_prior <- within(list(), {
  mu_mean <- rep(res_mex_war_summary["mu[1]", "mean"], 2)
  mu_cov <- diag(res_mex_war_summary["mu[1]", "sd"] ^ 2, 2)
  sigma_mean <- rep(res_mex_war_summary["sigma[1]", "mean"], 2)
  sigma_var <- rep(res_mex_war_summary["sigma[1]", "sd"] ^ 2, 2)
  rho_a <- res_mex_war_summary_rho$a
  rho_b <- res_mex_war_summary_rho$b
})

```

Run the model on the entire dataset to get initial values



Vague prior
```{r}
vague_prior <- within(list(), {
  mu_mean <- apply(y, 2, mean)
  mu_cov <- diag(rep(1e6, 2))
  sigma_mean <- apply(y, 2, sd)
  sigma_var <- rep(1e6, 2)
  rho_a <- 1
  rho_b <- 1
})
```

Casualty data to use. Transform casualties to log values (+1 to account of weeks with 0 casualties).

```{r}
y <- log(as.matrix(casualties_week[ , c("union", "confederate")]) + 1)
n <- nrow(y)
```

```{r}
m1_data_vague_all <- vague_prior
m1_data_vague_all$y <- y
m1_data_vague_all$n <- n
m1_results_vague_all <- sampling(m1, data = m1_data_vague_all, chains = 4)
init <- apply(m1_results_vague_all, 2, relist, 
              skeleton = rstan:::create_skeleton(m1_results_vague_all@model_pars, m1_results_vague_all@par_dims))

```

```{r}
run_model_subset <- function(i, prior, data, mod) {
  prior$y <- data[seq_len(i), ]
  prior$n <- i
  res <- sampling(mod, data = prior, verbose = FALSE)
  tmp <- summary(res)[[1]]
  tmp <- as.data.frame(tmp)
  tmp$parameter <- rownames(tmp)
  tmp$i <- i
  tmp
}


results <- vector(length = n, mode = "list")

results <- ldply(2:acw_data$n, run_model_subset,
                 prior = vague_prior, data = y, mod = m1)
results <- merge(results, mutate(casualties_week[ , "date", drop = FALSE], i = seq_along(date)))
save(results, file = "m1_vague.Rdata")

```


```{r}
(ggplot(filter(results, parameter %in% c("mu[1]", "mu[2]")) %>% mutate(parameter = factor(parameter)),
       aes(x = date, y = mean, ymin = mean - 2 * sd, ymax = mean + 2 * sd, colour = parameter))
 + geom_ribbon(alpha = 0.2) + geom_line())

```

```{r}
(ggplot(filter(foo, parameter %in% c("mu[1]", "mu[2]")) %>% mutate(parameter = factor(parameter)),
       aes(x = i, y = sd, colour = parameter))
 + geom_line())

```

```{r}
(ggplot(filter(results, parameter %in% c("sigma[1]", "sigma[2]")) %>% mutate(parameter = factor(parameter)),
       aes(x = date, y = mean, ymin = mean - 2 * sd, ymax = mean + 2 * sd, colour = parameter))
 + geom_ribbon(alpha = 0.2) + geom_line())

```

```{r}
(ggplot(filter(results, parameter %in% c("sigma[1]", "sigma[2]")) %>% mutate(parameter = factor(parameter)),
       aes(x = date, y = sd, colour = parameter))
 + geom_line())

```

# Mexican American War Prior

```{r}
results <- plyr::ldply(2:acw_data$n, run_model_subset,
                 prior = mex_war_prior, data = y, mod = m1, .progress = "text")
results <- merge(results, mutate(casualties_week[ , "date", drop = FALSE], i = seq_along(date)))
save(results, file = "m1_mex.Rdata")
```

```{r}
(ggplot(filter(results, parameter %in% c("mu[1]", "mu[2]")) %>% mutate(parameter = factor(parameter)),
       aes(x = date, y = mean, ymin = mean - 2 * sd, ymax = mean + 2 * sd, colour = parameter))
 + geom_ribbon(alpha = 0.2) + geom_line())

```

```{r}
(ggplot(filter(foo, parameter %in% c("mu[1]", "mu[2]")) %>% mutate(parameter = factor(parameter)),
       aes(x = i, y = sd, colour = parameter))
 + geom_line())

```

```{r}
(ggplot(filter(results, parameter %in% c("sigma[1]", "sigma[2]")) %>% mutate(parameter = factor(parameter)),
       aes(x = date, y = mean, ymin = mean - 2 * sd, ymax = mean + 2 * sd, colour = parameter))
 + geom_ribbon(alpha = 0.2) + geom_line())

```

```{r}
(ggplot(filter(results, parameter %in% c("sigma[1]", "sigma[2]")) %>% mutate(parameter = factor(parameter)),
       aes(x = date, y = sd, colour = parameter))
 + geom_line())

```

## Varying Data

Time varying model
```{r}
  m2 <- stan_model("stan/model2.stan")
```


```{r}
vague_prior_2 <-
  within(list(), {
      mu_mean <- rep(1, 2)
      mu_cov <- diag(rep(4, 2))
      sigma_mean <- rep(0.1, 2)
      sigma_var <- rep(2, 2)
      tau_mean <- rep(0.1, 2)
      tau_var <- rep(2, 2)
      rho_a <- 1
      rho_b <- 1
  })

```

```{r}
data_vague_prior_2 <- vague_prior_2
data_vague_prior_2$y <- y
data_vague_prior_2$n <- n

results_m2_vague_all <- sampling(m2, data = data_vague_prior_2, chains=1)
```
